---
title: "AT4BR_Exercise_6_Kinga_Zajdel_2024-04-016"
format: html
editor: visual
---

### 1 - Load the “iris” dataset.

```{r}
install.packages("tidyverse")
library (tidyverse)
iris_tb <- as_tibble (iris)
iris_tb
```

### 2 - Transform it to longer (tip: you should get 3 columns Species \| Trait \| Size )

```{r}
iris_tb_long<- iris_tb %>% 
  tidyr::pivot_longer(cols = !Species, names_to = "Trait", values_to = "Size") 

iris_tb_long
```

### 3 - Answer the following questions. (tip: you can answer each question “piping” only 2-3 commands )

-   A What is the **mean** length and width of the sepal and petal of each species?

```{r}
#iris_tb_mean<-iris_tb %>% 
#  dplyr::select("Sepal.Length", "Sepal.Width", "Petal.Length", "Petal.Width", #"Species", )
#
# rmarkdown::paged_table(iris_tb_mean) # shows the table

iris_tb_long  %>%  # using the long version with its cols!
  dplyr::group_by(Species, Trait) %>% 
  summarise(mean = mean(Size)) # assigning name of the column
```

-   B What is the **largest** and **widest**,  **petal** and **sepal**,  of **virginica** and **versicolor** species?

```{r}
iris_tb_long %>%  # using the long version with its col names!
  filter(Species %in% c("virginica", "versicolor")) %>% # no setosa
  group_by(Species, Trait) %>% # as above: grouping by two categories
  summarise(maximum = max(Size))
# From Console:
# `summarise()` has grouped output by 'Species'. You can override using the `.groups` argument.
```

-   C For each species, for **sepal** and **petal**, **width** and **length**,  how many observations are above the mean?*
(tip: adding the means as column might help)
*3C Might seem difficult (it is not that difficult if you know the tricks ;) ). If you don’t manage to make it work, submit your best attempts and we will discuss it in class.

```{r}

```

### A helpful webpage:
https://dplyr.tidyverse.org/articles

*No AI used.*
